---
title: "midterm_key_oct_8"
author: "Rob Wells"
date: "2025-10-15"
output: html_document
Owen Korzak Midterm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# JOUR405: Statistics for Journalists
## Midterm Exam - Fall 2025

For this exam, you'll analyze several datasets using R and the statistical concepts we've covered in class. Load the tidyverse before beginning, then complete each task. Write your code in the provided blocks and answer the questions in complete sentences.

# Load tidyverse
```{r}
library(tidyverse)
```

## Part 1: Restaurant Health Inspections (25 points)

The first dataset contains restaurant health inspection scores from Montgomery County. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv`

The dataset includes the name of the establishment, the number of points for critical and non-critical areas, the total points, maximum points possible and the compliance score and grade.

## Task 1: Read the data**
```{r}
data <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
```


## Task 2: Calculate the mean and standard deviation maximum and minimum of compliance scores. Put the results in a new dataframe called "compliance_stats." 
```{r}
compliance_stats <- data |>
summarize(mean_compliance = mean(compliance_score),
min_compliance = min(compliance_score) ,
max_compliance = max(compliance_score) ,
sd_compliance = sd(compliance_score)
)

```

## Task 3: Create a histogram of the compliance scores with a vertical line showing the mean. Include a headline. Include a headline and label the x and y axes.
```{r}
data|> 
  ggplot() + 
  geom_histogram(aes(x = compliance_score), binwidth = 5) +
  geom_vline(aes(xintercept = mean(compliance_score)), color = "red", linetype = "dashed", size = 1) +
  labs(title="Compliance Scores for Restaurant Health Inspections")

```

## Task 4: Write 2-3 sentences interpreting what the standard deviation and histogram tell us about the distribution of compliance scores. What would be newsworthy about this distribution? What's the story here? 

**Write your answer here:**
In my opinion, the story would be about the restaurants that were below 90 on the compliance score. The vast majority of restaurants scored a 100, so writing about the restauarants that score well below the average (the min is 42!!) would be an interesting story, in my opinion.

## Part 2: High School Athletics (30 points)

The second dataset shows participation numbers in high school sports across Maryland counties in 2024, broken down by sport and gender. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv`

# Task 5: Load the high school participation data
```{r}
hs_sport_participation <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv")
```


#Task 6: Add two columns called total and girls_pct.  
Add two columns called total and girls_pct using mutate(), with the total adding together boys and girls and girls_pct being the percentage of the total represented by girls participants. 
```{r}
in_depth_hs_sport_participation <- hs_sport_participation|>
  mutate(total =  rowSums(across(3:4))) |>
  mutate(girls_pct = girls / total * 100)
```

#Task 7: Calculate correlation between boys and girls participation
```{r}
in_depth_hs_sport_participation |> summarize(correlation = cor(boys, girls, method="pearson"))
```

#Task 8: Create a scatterplot showing this relationship, adding a line of best fit. Include a headline and label the x and y axes.
```{r}
in_depth_hs_sport_participation |> 
  ggplot() +
  geom_point(aes(x=boys, y=girls)) +
  geom_smooth(aes(x=boys, y=girls), method="lm")+
  labs(title="Correlation between boys and girls HS sports partcipation")

```


# Task 9: In 2-3 sentences, explain what the correlation coefficient and scatterplot reveal about gender equity in Maryland high school sports participation. How do you interpret the school districts that are below the line vs those that are above? Which school districts are most worth examining further, and why? (10 points)

**Write your answer here:**
There is a very strong correlation between between boys and girls HS sports participation. The correlation coefficient is 982419, which is extremely close to 1, which would be a direct correlation. No outliers pop out. If a district is above the line of best fit, then that means there are more girls participating in sports than expected. If a district is below, then it means there are less boys participating than expected. The ones worth examining more are the ones that are very far away from the line of best fit, which most are fielding less boys than expected.

## Part 3: Maryland Car Theft Rates (25 points)

Your editor has assigned you a story about car thefts in Maryland and wants you to analyze the data to find out which counties have the highest rates. The fourth dataset contains car theft statistics for Maryland counties in 2023 and population. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv` and complete the following tasks:

#Task 10: Load the car thefts data
```{r}
car_thefts_data <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv")
```

#Task 11: Using mutate, add a column that calculates the rate of car thefts for each county - you need to choose the per capita rate 
```{r}
car_thefts_data_rate <- car_thefts_data|>
  mutate(car_theft_rate = (2023 / population) * 1000)
```

#Task 12: Calculate the median car theft rate and the total number of car thefts statewide. Which counties have rates above the median? 
```{r}
car_thefts_data_rate |>
  summarize(median = median(car_theft_rate))

car_thefts_data_rate |>
  summarize(total =  colSums(across(4:4)))

```
Total amount of car thefts: 23,871
Counties above the median: Allegany, Calvert, Caroline, Cecil, Dorchester, Garrett, Kent, Queen Anne's, St. Mary's, Talbot, Wicomico, Worchester

#Task 13: Write 2-3 sentences describing what these calculations reveal about the distribution of car thefts in Maryland. What's the lede of a story about your findings? 

**Write your answer here:**

Something I found in the data is that, despite the more populated areas having high numbers of car thefts, their rate, per capita of 1000, is very low. I think that a good story is that the populated cities where many outsiders think everyone may be a criminal (Baltimore pretty much), actually has a very low car theft rate, per 1000 people.

## Part 4: Data Analysis Scenario (20 points)

#Task 14: Essay Question.

You receive a tip that local emergency response times have gotten significantly worse over the past year. You obtain monthly data on response times for police, fire and ambulance calls.

Write 3-4 sentences explaining:
1. What statistical measures would you calculate to verify this claim? (10 points)
2. What visualizations would help readers understand the trends? (5 points)
3. What additional context or data would you need to make this a complete story? (5 points)

**Write your answer here:**

# Excellent Response

The statistical measure I would calculate to verify this is average response time. I would also incorporate the months. I would probably use a histogram, the x axis would be month, the y axis would be average response time. It would be clear to the audience to see if there is a trend in how quick the response time is. 
Additional context I would need is likely how many calls they are getting. If the amount of calls has gone up, it is fair to assume that it may take longer for a response.

### Submission Instructions
- Save your work frequently
- Make sure all code blocks run without errors
- Provide clear explanations for your analytical choices
- Submit both your code and written responses
- Load this test to GitHub and submit the link in Elms

Good luck!
