---
title: "Oct 1 Activities: M&Ms; Survey; Sampling"
date: 9-30-2025
---

All three class activities are on this .rmd notebook

Activity #1: Activity: M&Ms - Central Limit Theorem

Working in groups, each group will get a handful of candy from the bag, and then count the number of red and black candies. Log those in this Google Sheet 

https://docs.google.com/spreadsheets/d/1KOeoVTq2W7L2T9sWjwh5U_2f59GsTeAylVGD53rNWww/edit?usp=sharing

(give your group a nice name).

Take a look at the values that we all get - what do you notice about the patterns? This is an exercise about the central limit theorem. 

The central limit theorem states that the sampling distribution of the mean of any independent, random variable will be normal or nearly normal, if the sample size is large enough.

Then discuss and try to answer these questions:

--How close is your group's mean to the true proportion of 0.5?
--Why do different groups get different results?
--What happens to individual extreme results when we combine all groups?
--How does this relate to polls and surveys in real life?

Write your answer below. And then you may eat the M&Ms.

**A:**
The patterns our group found that each color m&m was pretty evenly distributed.

I think that if this sample was done 100s of times, then the means would all be pretty similar.


Our groups mean was not close to half at all. The colored M+Ms dominated the proportion, with 89%. Most groups had similar numbers to us. I guess this can be expected when you are calculating one color vs the other 5.  However, each color looks pretty similarly distributed. 14% of the M&ms are red, coming a lot closer to the 11% that are brown.  If the results are combined, then the average for a bag of m&ms would be a lot more accurate than looking at just one group, about 1/3 of the bag.  This relates to real life, as sampling is more accurate when you look at a much large sample. 
------------------------------------------------------------------------------

# Activity #2:

Belonging Survey

Working in pairs, take a look back at the main findings from UMD's Belonging Survey:

https://president.umd.edu/internal/belonging


Specifically, focus on the "Sample Description" section and the Appendix. Discuss the differences and between the population and the sample, and what the document says about how the survey attempted to deal with them. Are there any decisions described in the document that raise questions in your minds? Detail the differences you see and your understanding of the methods they employed (it's ok if you don't understand all of them), and add your questions. You should discuss this as a pair but submit individually.

**A:**
Only 22.8% of the UMD community actually responded to the belonging survey, meaning the data does not necessarily reflect how the entire campus community feels. According to the appendix, "Over 25,000 qualitative responses were analyzed using content analysis to extract common themes and thematic patterning across respondents."  This is how the University was able to deal with differences. Something that raises questions for me is how the university handles people in the sample who don't identify as male or female. I wonder if people who don't identify or people who don't specify have a different impact on the survey, since they are not as big of a group.
------------------------------------------------------------------------------

# TASK 3: SAMPLING

Based on the class sampling demo (available on Elms), sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

**A**
```{r}
library(tidyverse)

```


### Task 2: Load 1,000 rows of data with the ages of people from 18-80

**A**
```{r}
set.seed(42)
ages <- tibble(
  age = sample(18:80, 1000, replace = TRUE)
)

head(ages)

```

### Task 3: Sample 100 rows of data from the dataset
**A**
```{r}
sample_50 <-ages |> 
  sample_n(50)

```


### Task 4: Calculate the mean age and standard deviation of the sample
**A**
```{r}
sample_50 |> summarize(mean = mean(age), sd = sd(age))


```


### Task 5: Calculate the mean age and standard deviation of the population
**A**
```{r}
ages |> summarize(mean = mean(age), sd = sd(age))

```


### Task 6: Compare the sample and population statistics
**A**

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

**A**: The means and standard deviations are very close.  The population has a higher mean but a lower standard deviation.  In a smaller sample, the mean and standard devation are both lower than the population, but again, the numbers are still close.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
