---
title: "Data Distribution, Volleyball Percentiles, Car Thefts, "
author: "Derek Willis /Rob Wells"
date: 09-23-2025
---

This workbook contains three exercises for our class. Complete them and hand it in at the end of class via a GitHub link.

Your name:
**Answer**
Owen Korzak
#------------------------------------------------------------------------------#


# Part 1: When Data Isn't Normal

Working in pairs, consider the following question: "Why do many statistical methods assume a normal distribution, and what challenges arise when data does not meet this assumption?"

First, come up with two examples of data that might not be normally distributed - in other words, that could be skewed or unequal in some way - and explain why they wouldn't be normally distributed. Feel free to draw what you think the distribution would look like.

Then, how should a reporter working on a story about data that isn't normally distributed describe that data? Which statistical measures that we've covered in class should they use or avoid when doing so? Use the examples you came up with, and you should discuss your answers together, but each student submits separately.

**Your answer**:
One instance where data may not be normally distributed is income.  I would assume that the tail of the distribution would be on the right, as not as many people in the US, for example, make more than, lets say, $150,000. Another instance could be age. For example, a country like Japan, which has an older population, will have it's tail on the left side of the data.

I think something a reporter should do if data is not normally distributed is making sure that when they are looking for sources, they don't do it by convenience. Additional research may be necessary. In the olympic example we talked about in class, a good way to analyze the data could be by looking at data like the amount of money spent per athlete.


#------------------------------------------------------------------------------#


# Part 2: Volleyball Percentiles

We'll be calculating percentiles to see how good Maryland was in service aces in the 2024 season. We'll use `mutate()` and a new function `ntile()` to calculate percentiles. Get started by loading the tidyverse in a code block, then load the CSV file here: https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/refs/heads/main/data/ncaa_womens_volleyball_matchstats_2024.csv" and save it to a dataframe called `matches`.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the data

```{r}
matches <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/refs/heads/main/data/ncaa_womens_volleyball_matchstats_2024.csv")
```

### Task 3: Create totals of aces for each team using `group_by()` and `summarize()`

```{r}
team_aces <- matches |> 
  group_by(team) |> 
  summarize(total_aces = sum(aces)) |> 
  arrange(desc(total_aces))
```


### Task 4: Calculate percentiles using `mutate()` and `ntile()`

```{r}
team_aces |> mutate(percentile = ntile(total_aces, 100))
```


### Task 5: Describe the percentiles

Write a couple of sentences that puts Maryland into context with the other teams in the dataset. You can do some Internet research to help with this.

**Answer**:
Maryland is in the 95th percentile.  This means that Maryland has more or the same amount of aces as 95% of the teams in the data set.  This means Maryland is really good, compared to the rest of the teams in the data set, when it comes to getting aces. Aces are when you serve the ball and your opponent is unable to return the ball back to the service team. Aces seem to not be an indicator of whether or not a volleyball team is good. Penn State, the 2024 national champion, is ranked #41 nationwide in aces.

#------------------------------------------------------------------------------#

# Part 3: Car Thefts

Follow along with the demo to learn how to calculate standard deviation and how to create a histogram of the data using R and the Tidyverse. Get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```


### Task 2: Load the car thefts dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/car_thefts_jan24.csv


```{r}
car_thefts <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/car_thefts_jan24.csv")
```

### Task 3: Calculate the standard deviation of the number of car thefts, along with the mean

```{r}
car_thefts |> summarize(mean = mean(count), sd = sd(count))
```


### Task 4: Create a histogram of the number of car thefts


```{r}
car_thefts |> 
  ggplot() + 
  geom_histogram(aes(x = count), binwidth = 2)
```




### Task 5: Add a line to the histogram that shows the mean of the number of car thefts, and two other lines to show one standard deviation in both directions

```{r}
car_thefts |> 
  ggplot() + 
  geom_histogram(aes(x = count), binwidth = 2) +
  geom_vline(aes(xintercept = mean(count)), color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(count) - sd(count)), color = "blue", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(count) + sd(count)), color = "blue", linetype = "dashed", size = 1)
```

### Task 6: Describe your results

Is the distribution normal or skewed in one direction? Write a sentence in which you describe the range of the data covered by one standard deviation from the mean on both sides - how much of the data falls in that range?


**Answer**:

The median is larger than the mean, which means that the data is skewed to the left. 68.2% of the data is within one standard deviation of the mean. 34.1% of the data is within the red line to one of the blue lines.


When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.

Congrats!


