---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities |>
  mutate(violent_rate_2019 = (violent2019 / pop2019) * 1000)|>
  mutate(property_rate_2019 = (property2019 / pop2019) * 1000)|>
  mutate(violent_rate_2020 = (violent2020 / pop2020) * 1000)|>
  mutate(property_rate_2020 = (property2020 / pop2020) * 1000)
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

ANSWER HERE
The first thing that sticks out is that both rates dropped from 2019 to 2020. This is more than likely due to COVID-19. Dorchester and Baltimore City seem to be the leaders of each category. Crime in Maryland is likely highly swayed by those cities, as they are much more populated compared to the rest of the state.

### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |>
  mutate(violent_pct_change = ((violent_rate_2020 - violent_rate_2019) / violent_rate_2019) * 100)|>
  mutate(property_pct_change = ((property_rate_2020 - property_rate_2019) / property_rate_2019) * 100)

```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

ANSWER HERE

Like I said, most of the change is likely due to COVID; most of the cities had negative changes from 2019 to 2020.  A few cities had some large, positive changes, like Rising Sun and Seat Pleasant. This is likely an outlier and would probably dig a little bit deeper into this.


### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

ANSWER HERE
Lede: COVID-19 Impacts on Property and Crime Rates in Maryland.

Nutgraf:
COVID-19 kept many stuck at home. As a result, crime rates began to diminish all over the state of Maryland. However, property rates didn't take as much of a hit. As people were forced to stay home, many may have looked at what other housing options were available. As a result, some areas in Maryland took a leap in property rates.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
