---
title: "HW6_Spotify_Songs"
---

In this assignment, you will be working with a dataset of songs played on Spotify in 2020, and you'll be writing code to explore the range of danceability scores in the data, using standard deviation and histograms. You'll also write up your findings.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the Spotify Songs dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv

```{r}
spotify_danceability <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv")

read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv")
```

### Task 3: Calculate the mean and standard deviation of the danceability scores

Write code to calculate the mean and standard deviation of the danceability scores in the Spotify dataset. Display those.

```{r}
spotify_danceability |> summarize(mean = mean(danceability), sd = sd(danceability))
```

### Task 4: Create a histogram of the danceability scores

Write code to create a histogram of the danceability scores in the Spotify dataset, based on the example we did in class.

```{r}
spotify_danceability |> 
  ggplot() + 
  geom_histogram(aes(x = danceability), binwidth = .05)
```

### Task 5: Add a line to the histogram for the mean danceability score

Take the same code and add a vertical line to the histogram for the mean danceability score, based on the example in class.

```{r}
spotify_danceability |> 
  ggplot() + 
  geom_histogram(aes(x = danceability), binwidth = .05) +
  geom_vline(aes(xintercept = mean(danceability)), color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(danceability) - sd(danceability)), color = "blue", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(danceability) + sd(danceability)), color = "blue", linetype = "dashed", size = 1)
```
I accidentally did both the mean and SD here beacuse I was going off what we did in class. My bad.

### Task 6: Add two lines to the histogram to show one standard deviation

Take the previous code and add two lines to the histogram for the 1st standard deviation below and above the mean score.

```{r}
spotify_danceability |> 
  ggplot() + 
  geom_histogram(aes(x = danceability), binwidth = .05) +
  geom_vline(aes(xintercept = mean(danceability)), color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(danceability) - sd(danceability)), color = "blue", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(danceability) + sd(danceability)), color = "blue", linetype = "dashed", size = 1)
```

### Task 7: Describe our results and pitch a story

Write a few sentences about the distribution of danceability scores in the Spotify dataset. Is it a normal distribution, or is it skewed in one direction? What does the mean and one standard distribution tell us about the songs in the dataset? Draft a short story pitch from this data.

The data shows us that the majority of the songs have a  danceability score higher than 50%, which is good.The data is skewed to the left. One standard deviation contains around 68% of the songs.  A story could be thata very high quantity of the songs in the data have "good" danceability score, as the mean is a realatively high number.  

### Task 8: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
