---
title: "HW7_MandMs"
date: 10/1/2025
---


In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```


### Task 2: Load an M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
mandms <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```


### Task 3: Sample 5 rows of data from the dataset, and then sample 10 rows, saving each to separate dataframes called sample5 and sample10. Be sure to use set.seed(1234) before you start.
```{r}
set.seed(1234)

sample5 <- mandms |> 
  sample_n(5)

sample10 <- mandms |> 
  sample_n(10)

```


### Task 4: Calculate the mean, median and standard deviation of red M&Ms from each of the samples.
```{r}
sample5 |> 
  summarise(mean = mean(red), median = median(red), sd = sd(red))

sample10 |> 
  summarise(mean = mean(red), median = median(red), sd = sd(red))
```


### Task 5: Calculate the mean, median and standard deviation of red M&Ms from the population

```{r}
mandms |> 
  summarise(mean = mean(red), median = median(red), sd = sd(red))
```

### Task 6: Make three histograms: one each for sample5, sample10 and the original dataset. Include a headline. Use a consistent binwidth.

```{r}
ggplot(sample5, aes(x=red)) +
  geom_histogram(binwidth=1, fill="red", color="black") +
  labs(title="Histogram of M&Ms (Sample of 5)")


ggplot(sample10, aes(x=red)) +
  geom_histogram(binwidth=1, fill="red", color="black") +
  labs(title="Histogram of M&Ms (Sample of 10)")


ggplot(mandms, aes(x=red)) +
  geom_histogram(binwidth=1, fill="red", color="black") +
  labs(title="Histogram of Red M&Ms (Population)")
```


### Task 7: Answer questions


Compare the three sets of means, standard deviations and histograms. Which are more helpful to you in understanding the accuracy of the samples?

The population gives the true distribution. The sample of 10 looks closer to the population than the sample of 5. Larger samples usually reflect the population better.

What does this exercise tell you about the relationship between sample size and estimate accuracy?

Larger samples give more accurate estimates because they reduce random variation.

How does random sampling introduce variation into statistical estimates?

Each random draw gives slightly different numbers for mean, median, and standard deviation.

If you repeated this process, would the samples always be identical? Why or why not?

No, they would differ because random sampling produces different subsets each time.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
